### 2020-04-11 paper/paper.tex
intro and overview

%% This is an abbreviated template from http://www.sigplan.org/Resources/Author/.

\documentclass[acmsmall,review,authorversion]{acmart}
\acmDOI{}
% \acmJournal{FACMP}
\acmVolume{CSCI 5535}
\acmNumber{Spring 2020}

\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{The Name of the Title is Hope}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{TBD}
\email{tbd@colorado.edu}
\author{TBD}
\email{tbd@colorado.edu}
\affiliation{%
  \institution{University of Colorado Boulder}
}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  TBD
\end{abstract}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}

One of the more powerful tools in use by functional programmers are engines for
searching over code bases given the type signature of the desired function. Such
tools provide the entry-point for practical code reuse and management of large
code bases. Tools such as Hoogle(for Haskell) provide a best-in-class search
functionality to polymorphic functional languages, but the same tools are not as
prevalent for dependently-typed languages.

There are a number of factors that have prevented robust search features from
entering the dependently-typed landscape. The primary setback is the
undecidability of type equality(isomorphism) in the presence of dependent types.
For search over types to be performed there must be some way of saying whether
two types are equal or not. This introduces many of the nuances of equality into
the confusion of trying to build a practical tool.

We attempt to solve this problem by defining an abstraction which defines a
search engine for dependent types, that can scale and adapt to appropriate
definitions of type equality. It is reasonable for a user to want some control
over how weak/strong the equalities for search are. Some may want strong
definitional equalities, others may accept the more computationally burdensome(and
semi-decidable) isomorphisms. We give an approach that leaves this control to
the user, and also leaves open the possibility for very powerful forms of
program search.

Another deficiency of current search engines for program source is the lack of
being able to specify complex functional properties over the domain of search.
Input-Output examples and algebraic properties offer two great sources of
specification for search, yet they are not used, typically because of the lack
of some method for verifying these properties. Fortunately, a dependently
typed language offers a natural way of handling these properties.

Practically, we provide all of the abstractions formalized in Agda and presented
as concrete tool for search over functions with input-output examples. The
implementation is capable of being extended to support more expressive
properties which in turn will provide more power to a user to specify complex
queries.

\section{Overview}

Let's start with an example of a search query over a non-dependent type, given
as a type:

$$
\text{query} : \mathbb{N} \rightarrow \mathbb{N} \rightarrow \mathbb{N}
$$

Solutions for this query would obviously include the usual arithmetic
operations. But let's say that we wanted to further specify the query by
providing some input-output examples:

$$
\begin{array}{ccc}
  \text{input 1} & \text{input 2} & \text{output}\\
  0 & 0 & 0\\
  2 & 2 & 4\\
  1 & 2 & 3\\
  ... &...&...
\end{array}
$$

We could devise the obvious procedure of checking the resulting functions
against these examples and discarding those that are not consistent. In fact,
this is exactly what we will do. But to jump to the algorithm would fail to
realize the deeper type-theoretic possibilities here. See, the examples above
can be encoded quite easily as a type for our query. The type below is
consistent with the query:

$$
\begin{array}{c}
\Sigma[\text{f} \in \mathbb{N} \rightarrow \mathbb{N} \rightarrow
  \mathbb{N}] ((f\ 0\ 0\ \equiv 0) \times (f\ 2\ 2\ \equiv 4) \times (f\ 1\ 2\
  \equiv 3))
\end{array}
$$

This is a \textit{dependent product} type that encodes our query as searching
for a term that satisfies the type we gave along with proofs that the provided
function produces certain results under evaluation.

\section{(Contribution 1)}

\section{(Contribution 2)}

\section{Empirical Evaluation}

\section{Related Work}

\section{Conclusion}

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
TBD
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}
\end{document}

### 2020-04-25 paper/paper.tex
technical sections

@@
 
+\usepackage{xargs}                      % Use more than one optional parameter in a new commands
+\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
+\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
+\newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
+% \newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
+% \newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
+% \newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
+
+\newcommand{\NC}{(\textbf{Needs Citation})}
+
+\newcommand{\term}[1]{\textsf{#1}}
+\newcommand{\type}[1]{\textsf{#1}}
+
+\newcommand{\Carrier}{\term{Carrier}}
+\newcommand{\Universe}{\mathcal{U}}
+\newcommand{\Set}{\type{Set}}
+\newcommand{\Type}{\type{Type}}
+\newcommand{\Boolean}{\term{Boolean}}
+\newcommand{\true}{\term{true}}
+\newcommand{\false}{\term{false}}
+\newcommand{\Number}{\term{Number}}
+\newcommand{\Interp}[1]{\llbracket #1 \rrbracket}
+\newcommand{\?}{\stackrel{?}{\approx}}
+\newcommand{\search}{\term{search}}
+
+\newcommand{\data}[2]{\textsf{data } #1 : #2 \textsf{ where}}
+\newcommand{\record}[2]{\textsf{record } #1 : #2 \textsf{ where}}
+
 \begin{document}
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{The Name of the Title is Hope}
+\title{Embedded Program Search in Agda}
 
 %%
-%% The "author" command and its associated commands are used to define
+%% The "author" command and its associa
+
 %% the authors and their affiliations.
@@
 %% used to denote shared contribution to the research.
-\author{TBD}
-\email{tbd@colorado.edu}
-\author{TBD}
+\author{Jack Martin}
+\email{John.P.Martin@colorado.edu}
+\author{Michael Dresser}
 \email{tbd@colorado.edu}
@@
 \begin{abstract}
-  TBD
+The problem of Program Search - looking for a function which satisfies a query -
+is one of the more practical tools a developer can use when programming. Search
+tools such as Hoogle provide an entrypoint for a language eco-system, but
+unfortunately often don't accept query information other than types. Here we
+present a search tool for searching over types and examples as well as a general
+framework for doing program search over dependent types. We also provide a
+type-theoretic interpretation of the program search problem as a special case of
+the more general synthesis problem and discuss how this presents a tool for
+furthering the solution to both problems.
 \end{abstract}
@@
 
-\section{(Contribution 1)}
+\section{Preliminaries}
 
-\section{(Contribution 2)}
+\todo[inline]{Note for Benno and Evan: The citations are absent for now. We will
+add these for the final paper but decided to leave them out for now until the
+structure of the paper is fixed. The notes for ``Needs citation'' indicate
+positions where we have a reference saved somewhere. It is not an indication
+that we need to still find it. }
+
+\subsection{Universes} One of the advantages of dependent type theories is their
+ability to formalize their own type system and meta-theories within
+themselves(\textbf{Needs citation}). These techniques typically use a structure
+called a \textit{Universe}, which is defined as a type along with an
+interpretation function which takes an element of that type and computes some
+type in the type system. Formally:
+
+$$
+\begin{array}{l}
+\record{\Universe}{\Set}\\
+\quad  \begin{array}{rl}
+         \Carrier &: \Set \\
+         \Interp{\_} &: \Carrier \rightarrow \Set
+       \end{array}
+\end{array}
+$$
+
+An example of such a universe containing booleans, natural numbers, and
+functions on either is given here:
+
+$$
+\begin{array}{cc}
+\begin{array}{l}
+\data{\Type}{\Set}\\
+\quad \begin{array}{rl}
+        \textsf{Number} :& \Type\\
+        \textsf{Boolean} :& \Type\\
+        \_\Rightarrow\_ :& \Type \rightarrow \Type \rightarrow \Type
+      \end{array}
+\end{array}
+&
+\begin{array}{l}
+  \Interp{\_}_{\Type} : \Type \rightarrow \Set\\
+  \Interp{\textsf{Number}}_{\Type}     = \mathbb{N}\\
+  \Interp{\textsf{Boolean}}_{\Type}    = \mathbb{B}\\
+  \Interp{t_1 \Rightarrow t_2}_{\Type} \ \, = \Interp{t_1} \rightarrow \Interp{t_2}
+\end{array}
+\end{array}
+$$
+
+Soon, we will define a universe of types for use in program search. We will
+spend most of the rest of the paper defining a suitable $\Carrier$ type that
+respects the various complexities that arise from the presence of dependent
+types.
+
+\subsection{Simulating First Class Modules}
+
+We will need a coarse simulation of Agda's module structure to store entries for
+search. This will be accomplished by collecting elements of a $\Carrier$ type
+along with the terms they encode into a homogeneous list. The first element will
+be the coded type, and the second element(which depends on the first), will be
+the term of the type that is interpreted from the code:
+
+$$
+\textsf{Module} = \Sigma[ \tau \in \Carrier ] \Interp{ \tau }
+$$
+
+Carrying on with our example universe above, an example module containing some
+common functions and elements:
+
+$$
+\begin{array}{rl}
+  \textsf{example} =& (\Boolean\ ,\ \true)\\
+                  ::& (\Number\ ,\ 5 )\\
+                  ::& (\Boolean\ ,\ \false )\\
+                  ::& (\Number \Rightarrow \Boolean\ ,\ (== 5) )\\
+                  ::&[]
+\end{array}
+$$
+
+\section{Searching over Universe Modules}
+
+To support basic type search over these modules we will first have to discuss
+what it means for a query type to ``match'' a type in a collection. This clearly
+introduces the need for a notion of equality over types, which complicates
+matters quickly. Which form of equality of types is appropriate for use in
+program search? Fortunately, there has been some research done on this problem
+and the current accepted equality is isomorphism of types\NC. This
+will equate curried function types to their uncurried counterparts for
+example(i.e. $(A \times B) \rightarrow C \approx A \rightarrow B \rightarrow
+C$).
+
+A problem with using isomorphism over dependent types is that it is a generally
+undecidable problem\NC. The field is currently studying certain decidable
+subsets of the problem thus we opt to maximize flexibility by parameterizing our
+methods over a provided decidable type equality. The problem of type-directed
+program search then just becomes the act of running the decision procedure for
+this equality over the coded-types in our module. Thus, a searchable universe
+can be defined:
+\todo{define isEquiv}
+$$
+\begin{array}{l}
+  \record{\Universe_S}{\Set}\\
+  \quad \begin{array}{ll}
+          U &: \Universe\\
+         \_\approx\_ &: (a\ b\ : \Carrier\ U) \rightarrow \Set\\
+         \approx-\textsf{isEquiv} &: \textsf{IsEquivalence } (\Carrier\ U)\ \_\approx\_\\
+         \_\?\_ &: \forall (a\ b : \Carrier\ U) \rightarrow \textsf{Dec } (a \approx b)
+        \end{array}
+\end{array}
+$$
+
+Now we are ready to define an abstract search procedure for modules of a
+decidable universe. The type of the search function is:
+
+$$
+\begin{array}{l}
+  \search : (u : \Universe_S) \rightarrow (m : \textsf{Module}_u)\rightarrow (c : \Carrier\ (U\ u))
+  \rightarrow \textsf{Maybe }\Interp{c}
+\end{array}
+$$
+
+The procedure takes in a decidable universe $u$, a module of that universe $m$,
+and a coded-type $c$. The output type of the function is possibly an inhabitant
+of the interpretation of $c$ provided that there is in fact one in the module.
+
+\section{A Universe for Searching over Dependent Types}
+
+Thus far, the formalism we have given should allow any non-dependent program
+to be searched. To add support for programs with type dependencies we will need
+to slightly alter our definition of $\textsf{Module}$. The main idea is that the
+module will need to be simultaneously defined with the interpretation function
+so that types defined in a ``cons'' of a module can depend on the terms found in
+earlier entries.
+
+The definition of this universe is given in Figure \ref{fig1}.
+\todo[inline]{Note for Benno and Evan: There is not any explanation or elaboration for this
+  part of the formalism yet. There are also a few missing constructors. We were
+  stuck on this for the better part of the last week and just figured out how to
+formalize it last night and it isn't quite finished. In the final paper this
+will be much, much more spelled out over 3-5 pages.}
+
+\begin{figure}[t]
+  $$
+  \begin{array}{cc}
+  \begin{array}{l}
+    \data{\textsf{Module}}{\mathbb{N} \rightarrow \Set}\\
+    \quad \begin{array}{ll}
+            []        &: \textsf{Module } 0\\
+            \_::\_,\_ &: \forall \{ n \}\\
+                      &\rightarrow  (\Gamma : \textsf{Module } n)\\
+                      &\rightarrow  (t : \textsf{TypeCode } n)\\
+                      &\rightarrow  \Interp{ t }_\Gamma \\
+                      &\rightarrow \textsf{Module (suc }  n)
+          \end{array}
+  \end{array}
+                        &
+  \begin{array}{l}
+    \data{\textsf{TypeCode}}{\mathbb{N} \rightarrow \Set}\\
+    \quad
+    \begin{array}{ll}
+      U &: \forall \{ n \} \rightarrow \textsf{TypeCode } n\\
+      \# &: \forall \{ n\ m \} \rightarrow n < m \rightarrow \textsf{TypeCode } m\\
+      \_\Rightarrow\_ &: \forall \{ n \}\ (t_1\ t_2 : \textsf{TypeCode } n)
+                        \rightarrow \textsf{TypeCode } n
+    \end{array}
+
+  \end{array}
+  \end{array}
+  $$
+  $$
+  \begin{array}{l}
+    \Interp{\_}\_ : \forall \{ n \} \rightarrow \textsf{TypeCode } n \rightarrow
+    \textsf{Module } n \rightarrow \Set\\
+    \Interp{ U }_\Gamma = \Set\\
+    \Interp{ t_1 \Rightarrow t_2 }_\Gamma = \Interp{t_1}_\Gamma \rightarrow \Interp{t_2}_\Gamma\\
+    \Interp{ \#\ \textsf{suc-leq-suc ( zero-leq-n)}}_{(\Gamma :: U ,
+    \textsf{term})} = \textsf{term}\\
+    \Interp{ \#\ \textsf{suc-leq-suc ( zero-leq-n)}}_{(\Gamma :: t ,
+    \textsf{term})} = \langle\textsf{Error Case - Impossible under assumptions}\rangle\\
+    \Interp{ \#\ \textsf{suc-leq-suc ( suc-leq-suc } n )}_{(\Gamma :: t ,
+    \textsf{term})} = \Interp{ \#\ \textsf{suc-leq-suc } n }_\Gamma
+  \end{array}
+  $$
+  \caption{The Universe for Search over Dependent Types}
+  \label{fig1}
+\end{figure}
 
@@
 
+\todo[inline]{We were unsure what to include in this section because most of our
+  project was correct by construction. Benno and Evan, we will touch base with
+  you about your thoughts as to what to include here. We brainstormed a little
+  bit but couldn't settle on anything that made too much sense. Hoping that the
+  sections above will be suitable for this week's version of the draft. One idea
+  we had would be to show off an example of why the embedded program search is
+  so powerful/useful. This seems to be a somewhat empirical claim that we could
+  spend some time fleshing out.}
+
 \section{Related Work}
+
+\todo[inline]{Edit: Make changes as Benno suggested}
+
+\paragraph{Program Search} Program Search, or the search of functions in some
+library, has led to such tools as Hoogle, and is the main category that this
+work could be included under. In the past, none of these tools use any
+functional properties or examples to help guide the search. Ours is the first
+tool to provide this functionality.
+
+\textit{draft note: make sure to look into the search programs for other languages.}
+
+\paragraph{Program Synthesis} As was mentioned in Section \# above, program
+search can be viewed as a special case of the more general program synthesis
+problem. It seems hopeful that program search can be later used as a tool in
+program synthesis by treating it as an atomic operation for larger synthesis
+procedures.
+
+\paragraph{Type Isomorphism} Isomorphisms over types acts as the equivalence
+relation that we use for search in our approach. In reality our approach defines
+a slightly more robust isomorphism relation over types \textit{and} examples.
+This is slightly more powerful than just the types themselves. In future work we
+will look at the theoretical basis for this.
 

### 2020-05-06 src/helpers/ParserMain.hs
helper code from other repo

@@ -1,0 +1,14 @@
+module Main where
+
+import Data.List
+import System.Environment
+import System.Exit
+import Parser
+
+-- main :: IO ()
+main = getArgs >>=
+               \args -> (putStrLn "running with:") >>
+                        (putStrLn (show args)) >>
+                        (runParser (Data.List.head args)) >>
+                        return ()
+

### 2020-05-06 paper/paper.tex
big update

@@
 
+\newcommand{\N}{\mathbb{N}}
+\newcommand{\ra}{\rightarrow}
+
+\usepackage[utf8]{inputenc}
+\DeclareUnicodeCharacter{27E8}{$\langle$}
+\DeclareUnicodeCharacter{27E9}{$\rangle$}
+\DeclareUnicodeCharacter{2237}{$::$}
+\DeclareUnicodeCharacter{21D2}{$\Rightarrow$}
 \begin{document}
@@
 \author{Michael Dresser}
-\email{tbd@colorado.edu}
+\email{Michael.M.Dresser@colorado.edu}
 \affiliation{%
@@
 
-One of the more powerful tools in use by functional programmers are engines for
+Some of the more powerful tools in use by functional programmers are engines for
 searching over code bases given the type signature of the desired function. Such
 tools provide the entry-point for practical code reuse and management of large
-code bases. Tools such as Hoogle(for Haskell) provide a best-in-class search
-functionality to polymorphic functional languages, but the same tools are not as
-prevalent for dependently-typed languages.
+code bases. Tools such as Hoogle(for Haskell)\cite{mitchell2008hoogle} provide a
+best-in-class search functionality to polymorphic functional languages, but the
+same tools are not as prevalent for dependently-typed languages.
 
@@
 
-We attempt to solve this problem by defining an abstraction which defines a
-search engine for dependent types, that can scale and adapt to appropriate
+We attempt to solve this problem by defining an abstraction for
+search engines of dependent types that can scale and adapt to appropriate
 definitions of type equality. It is reasonable for a user to want some control
@@
 being able to specify complex functional properties over the domain of search.
-Input-Output examples and algebraic properties offer two great sources of
-specification for search, yet they are not used, typically because of the lack
-of some method for verifying these properties. Fortunately, a dependently
+Input-output examples and algebraic properties offer two great sources of
+specification for search, yet they are not used; typically this is because of
+the lack of methods for verifying these properties. Fortunately, a dependently
 typed language offers a natural way of handling these properties.
 
-Practically, we provide all of the abstractions formalized in Agda and presented
-as concrete tool for search over functions with input-output examples. The
-implementation is capable of being extended to support more expressive
-properties which in turn will provide more power to a user to specify complex
-queries.
+Practically, we provide all of the abstractions formalized in
+Agda\cite{norell2007towards} presented as a concrete tool for search over terms
+with some basic types. Notably, our method is embedded \textit{within} the type
+theory such that the search function can be treated just like any other
+function. The implementation is capable of being extended to support more
+expressive properties which in turn will provide more power to a user to specify
+complex queries. In the future, the theories in this paper will be developed
+into a search-directed approach to program synthesis.
 
@@
 
-Let's start with an example of a search query over a non-dependent type, given
-as a type:
+\subsection{Program Search}
 
-$$
-\text{query} : \mathbb{N} \rightarrow \mathbb{N} \rightarrow \mathbb{N}
-$$
+\newcommand{\extype}{\ensuremath \N \ra \N \ra \N}
 
-Solutions for this query would obviously include the usual arithmetic
-operations. But let's say that we wanted to further specify the query by
-providing some input-output examples:
+An advantage of the functional programming paradigm is the ability to compose
+programs from simpler functions. Large libraries exist containing many of the
+simple(and not-so-simple) functions that a programmer would like to avoid
+writing. However, the user of a functional language often runs into the
+situation where they have an idea of the function they need, but don't know the
+name or even where to begin looking. A robust tool for searching over the
+available libraries is the obvious solution.
+
+Let's assume we would like to search for a program that operates on two natural
+numbers. We have \texttt{plus} in mind, but can't remember the name of the
+function nor the module it may be located in. But we do know its behavior and
+type. The type of the function is $\extype$. We also can describe its behavior
+over a few input-output examples:
 
@@
 
-We could devise the obvious procedure of checking the resulting functions
-against these examples and discarding those that are not consistent. In fact,
-this is exactly what we will do. But to jump to the algorithm would fail to
-realize the deeper type-theoretic possibilities here. See, the examples above
-can be encoded quite easily as a type for our query. The type below is
-consistent with the query:
+The benefit of working in a dependent type theory is that we can encode these
+examples directly in our query type. So instead of $\extype$, we have the
+\textit{(iterated) dependent sum}:
+
+\begin{align*}
+\Sigma[ f \in \extype]\ &(f\ 0\ 0 \equiv 0)\\
+                         \times\ &(f\ 2\ 2 \equiv 4)\\
+                         \times\ &(f\ 1\ 2 \equiv 3)\\
+                         \times\ &...
+\end{align*}
+
+This type corresponds to the function $f$ of type $\extype$ along with proofs
+that the function satisfies our examples. A term of this type with \texttt{plus}
+as the function $f$ could look like:
+
+\begin{align*}
+  ( \texttt{plus}\ ,\ \texttt{refl}\ ,\ \texttt{refl}\ ,\ \texttt{refl}\ ,\ ...)
+\end{align*}
+
+Where \texttt{refl} is a reflexivity proof that a given example holds. The
+search for this term is quite simple. First, search for a term matching the type
+of the query, then check if the examples are satisfied by a reflexivity proof.
+Finding a term of the query type is fairly straightforward, but to find a proof
+that each example is held requires a bit more. Notably, the type of the output
+has a decidable (judgemental) equality. This is a property that would typically
+be proven in the same module as the types definition. Because of this, we can't
+make any assumptions about a given type possessing such a decidable equality. We
+should make the search procedure general enough that it can search for that
+decision procedure.
+
+This example is given to point out that the search methods we aim to define must
+be able to interact, to some degree, with the modules they search over. Simply
+looking for a matching type is not enough. For these reasons our search methods
+are embedded \textit{within} the type system as opposed to externally in the
+form of a tool or executable.
+
+\section{Preliminaries}
+
+\subsection{Dependently Typed Programming in Agda}
+
+This subsection briefly introduces the necessary ideas from dependently-typed
+programming necessary to understand the rest of the paper. Our paper will use
+a notation that roughly corresponds to that of the theorem prover Agda. We use
+datatype declaration of the form:
 
 $$
-\begin{array}{c}
-\Sigma[\text{f} \in \mathbb{N} \rightarrow \mathbb{N} \rightarrow
-  \mathbb{N}] ((f\ 0\ 0\ \equiv 0) \times (f\ 2\ 2\ \equiv 4) \times (f\ 1\ 2\
-  \equiv 3))
+\begin{array}{l}
+\data{T}{\tau_1 \ra \tau_2 \ra ... \ra \Set}\\
+\quad \begin{array}{rl}
+  \textsf{C}_1 & : T\ e_{1,1}\ e_{1,2}\ ...\\
+  \textsf{C}_1 & : \mathbb{N} \ra T\ e_{2,1}\ e_{2,2}\ ...\\
+        ... &\\
+  \textsf{C}_n & : T\ e_{n,1}\ e_{n,2}\ ...
+\end{array}
 \end{array}
@@
 
-This is a \textit{dependent product} type that encodes our query as searching
-for a term that satisfies the type we gave along with proofs that the provided
-function produces certain results under evaluation.
+Where $T$ is the name of the type(family) being declared. The types $\tau_i$
+represent the indices for a type family. $\Set$ is the type of types.
+$\textsf{C}_i$ are the constructors for the newly declared type and the
+right-hand side of the constructor definition is the type of the constructor
+along with the specific type in the family(specified with terms $e_{i,j}$ from
+the indice-types) to which it belongs.
 
-\section{Preliminaries}
+We specify iterated sums as record types in the style of Agda. These can be
+thought of as representing mathematical structures that are composed of several
+elements of different(potentially dependent) types. An example is the semigroup
+record:
 
-\todo[inline]{Note for Benno and Evan: The citations are absent for now. We will
-add these for the final paper but decided to leave them out for now until the
-structure of the paper is fixed. The notes for ``Needs citation'' indicate
-positions where we have a reference saved somewhere. It is not an indication
-that we need to still find it. }
+$$
+\begin{array}{l}
+ \record{\textsf{IsSemigroup } (A : \Set)}{\Set}\\
+  \quad
+  \begin{array}{rl}
+    \_\cdot\_ &: A \ra A \ra A\\
+    \textsf{assoc} &: \textsf{Associative } (\_\cdot\_)
+  \end{array}
+\end{array}
+$$
+
+This is parameterized over a type $A$ and defines a type containing a binary
+operation $\_\cdot\_$ along with a proof that $\_\cdot\_$ is associative. Note
+that the term $\textsf{assoc}$ is dependent on the term $\_\cdot\_$. The
+underscores in the semigroup operation denote positions for arguments to the
+operation. Agda is unique from most languages in that it supports this style of
+mixfix operators. As an example, applying $\_\cdot\_$ to arguments $a$ and $b$
+would be $a \cdot b$.
 
 \subsection{Universes} One of the advantages of dependent type theories is their
-ability to formalize their own type system and meta-theories within
-themselves(\textbf{Needs citation}). These techniques typically use a structure
-called a \textit{Universe}, which is defined as a type along with an
-interpretation function which takes an element of that type and computes some
-type in the type system. Formally:
+ability to formalize their own type system and meta-theories. These techniques
+typically use a structure called a \textit{Universe}, which is defined as a type
+along with an interpretation function which takes an element of that type and
+computes some type in the type system. Formally:
 
@@
 
-An example of such a universe containing booleans, natural numbers, and
-functions on either is given here:
+Here, \textsf{Carrier} is a type containing codes. These codes correspond to
+types in Agda's type system and $\Interp{\_}$ is the interpretation function,
+which takes an element of \textsf{Carrier} and computes its corresponding
+Agda-type. An example of such a universe containing booleans and natural numbers
+is given here:
 
@@
         \textsf{Boolean} :& \Type\\
-        \_\Rightarrow\_ :& \Type \rightarrow \Type \rightarrow \Type
       \end{array}
@@
   \Interp{\textsf{Boolean}}_{\Type}    = \mathbb{B}\\
-  \Interp{t_1 \Rightarrow t_2}_{\Type} \ \, = \Interp{t_1} \rightarrow \Interp{t_2}
 \end{array}
@@
 
-Soon, we will define a universe of types for use in program search. We will
-spend most of the rest of the paper defining a suitable $\Carrier$ type that
-respects the various complexities that arise from the presence of dependent
-types.
+We will spend most of the rest of the paper defining a suitable $\Carrier$ type
+for simulating the type system enough for type-based search.
+
+\section{Embedding Search}
 
@@
 $$
-\textsf{Module} = \Sigma[ \tau \in \Carrier ] \Interp{ \tau }
+\begin{array}{l}
+\textsf{Module } : \Set\\
+\textsf{Module} = \textsf{List } (\Sigma[ \tau \in \Carrier ]\ \Interp{ \tau })
+\end{array}
 $$
 
-Carrying on with our example universe above, an example module containing some
-common functions and elements:
+We can construct an example of a module using our example universe from the
+previous section:
 
@@
                   ::& (\Boolean\ ,\ \false )\\
-                  ::& (\Number \Rightarrow \Boolean\ ,\ (== 5) )\\
                   ::&[]
@@
 
-\section{Searching over Universe Modules}
+This new data structure will allow for us to work with terms along with their
+types as first class entities as long as we can encode them into this module
+structure.
+
+\subsection{Search over First Class Modules}
 
@@
 program search? Fortunately, there has been some research done on this problem
-and the current accepted equality is isomorphism of types\NC. This
-will equate curried function types to their uncurried counterparts for
-example(i.e. $(A \times B) \rightarrow C \approx A \rightarrow B \rightarrow
-C$).
+and the current accepted equality is isomorphism of
+types\cite{dicosmo2012isomorphisms}. This will equate curried function types to
+their uncurried counterparts. For example(i.e. $(A \times B) \rightarrow C
+\approx A \rightarrow B \rightarrow C$).
 
-A problem with using isomorphism over dependent types is that it is a generally
-undecidable problem\NC. The field is currently studying certain decidable
+A problem with using isomorphism over dependent types is its
+undecidability\cite{HoTT}. The field is currently studying certain decidable
 subsets of the problem thus we opt to maximize flexibility by parameterizing our
@@
 can be defined:
-\todo{define isEquiv}
+
 $$
@@
 
-Now we are ready to define an abstract search procedure for modules of a
-decidable universe. The type of the search function is:
+Where \textsf{IsEquivalence} is a proposition that $\_\approx\_$ is an
+equivalence over the \textsf{Carrier} type and $\Universe$ is the universe
+record-type defined above. Now we are ready to define an abstract search
+procedure for modules of a decidable universe. The type of the search function
+is:
 
@@
 \begin{array}{l}
-  \search : (u : \Universe_S) \rightarrow (m : \textsf{Module}_u)\rightarrow (c : \Carrier\ (U\ u))
-  \rightarrow \textsf{Maybe }\Interp{c}
+  \search : (m : \textsf{Module})\rightarrow (c : \Carrier) \rightarrow \textsf{List }\Interp{c}
 \end{array}
@@
 
-The procedure takes in a decidable universe $u$, a module of that universe $m$,
-and a coded-type $c$. The output type of the function is possibly an inhabitant
-of the interpretation of $c$ provided that there is in fact one in the module.
+The function takes in a module of the universe $m$, and a coded-type $c$. The
+output type of the function is a list of inhabitants of the interpretation of
+$c$. In the remainder of the paper we will construct a concrete Universe with
+which we will implement a small embedded search procedure.
 
-\section{A Universe for Searching over Dependent Types}
+\section{Searching over a Subset of Agda}
 
-Thus far, the formalism we have given should allow any non-dependent program
-to be searched. To add support for programs with type dependencies we will need
-to slightly alter our definition of $\textsf{Module}$. The main idea is that the
-module will need to be simultaneously defined with the interpretation function
-so that types defined in a ``cons'' of a module can depend on the terms found in
-earlier entries.
+To implement the formalism given in the previous section and create a practical
+search procedure, we begin by defining the universe of typing-codes
+corresponding to the \textsf{Carrier} field of the record type for universes.
+This definition is given in Figure \ref{fig:codes}. $U$ is the code for terms of
+type $\Set$. $\#$ is the reference type which is used to refer to a type that is
+defined elsewhere in the module, an example of its use will be shown below.
+$\_\Rightarrow\_$ is the code for function types. The natural number that
+indexes \textsf{TypeCode} is used for book-keeping purposes to ensure we don't
+ever over-index a module or create any circular references. It is a technical
+detail that we will not elaborate on.
 
-The definition of this universe is given in Figure \ref{fig1}.
-\todo[inline]{Note for Benno and Evan: There is not any explanation or elaboration for this
-  part of the formalism yet. There are also a few missing constructors. We were
-  stuck on this for the better part of the last week and just figured out how to
-formalize it last night and it isn't quite finished. In the final paper this
-will be much, much more spelled out over 3-5 pages.}
+Modules need to be redefined(See Figure \ref{fig:mods}) from the abstraction in
+the previous section because they allow for codes to reference other
+definitions. We will define Module similarly to the type synonym we used above,
+except that it will be mutually defined with the interpretation function. Note
+that for the $\_::\_$ case we provide a module first in the constructor so that
+the entry may depend on it. The entries in the module will be similar to that in
+the example module in Section 4. In the same way the dependent pair
+did in the previous Module definition.
+
+The interpretation function is given in Figure \ref{fig:interp}. As explained
+before, this function is responsible for mapping elements of $\textsf{TypeCode}$
+to their respective types in $\Set$. The cases are fairly straightforward except
+for references. The reference constructor $\#$ contains a proof that the index
+does not go over the length of the list. This is constrained by the indices
+mentioned before. The proof includes with it the position of the type we wish to
+reference and then the interpretation function looks it up via a recursive call.
+
+To make all of this a bit more concrete, Figure \ref{fig:ex} is an example of a
+module encoded into this typing universe. Here each entry is preceded by its
+type in the coded type system. Note that the references properly index the types
+of their functions or terms.
 
 \begin{figure}[t]
+  \centering
 $$  
-  \begin{array}{cc}
+  \begin{array}{l}
+    \data{\textsf{TypeCode}}{\mathbb{N} \rightarrow \Set}\\
+    \quad
+    \begin{array}{ll}
+      U &: \forall \{ n \} \rightarrow \textsf{TypeCode } n\\
+      \# &: \forall \{ n\ m \} \rightarrow n < m \rightarrow \textsf{TypeCode } m\\
+      \_\Rightarrow\_ &: \forall \{ n \}\ (t_1\ t_2 : \textsf{TypeCode } n)
+                        \rightarrow \textsf{TypeCode } n
+    \end{array}
+
+  \end{array}
+  $$
+  \caption{The Type Codes for our Universe supporting Search}
+  \label{fig:codes}
+\end{figure}
+
+\begin{figure}[b]
+  $$
   \begin{array}{l}
@@
   \end{array}
-                        &
-  \begin{array}{l}
-    \data{\textsf{TypeCode}}{\mathbb{N} \rightarrow \Set}\\
-    \quad
-    \begin{array}{ll}
-      U &: \forall \{ n \} \rightarrow \textsf{TypeCode } n\\
-      \# &: \forall \{ n\ m \} \rightarrow n < m \rightarrow \textsf{TypeCode } m\\
-      \_\Rightarrow\_ &: \forall \{ n \}\ (t_1\ t_2 : \textsf{TypeCode } n)
-                        \rightarrow \textsf{TypeCode } n
-    \end{array}
+$$
+\caption{Modules Re-Defined}
+\label{fig:mods}
+\end{figure}
 
-  \end{array}
-  \end{array}
-  $$
+  \begin{figure}
+    \centering
   $$
@@
   $$
-  \caption{The Universe for Search over Dependent Types}
-  \label{fig1}
+  \caption{The Interpretation Function}
+  \label{fig:interp}
 \end{figure}
 
-\section{Empirical Evaluation}
+\begin{figure}
+  \centering
+\begin{verbatim}
+  ex : Module _
+  ex = [] ∷ U             , Nat
+          ∷ ⟨ 1 ⟩         , 5
+          ∷ ⟨ 1 ⟩         , 2
+          ∷ U             , Bool
+          ∷ ⟨ 4 ⟩         , true
+          ∷ ⟨ 4 ⟩         , false
+          ∷ ⟨ 4 ⟩ ⇒ ⟨ 4 ⟩ , not
+\end{verbatim}
+  \caption{A Sample Module.
+  The $\langle\_\rangle$ used in this example corresponds to a shorthand for
+  generating the safety proof automatically instead of writing it out long-hand
+  with $\#$. These can be read as references where the number inside the
+  brackets is the location of the type to which it refers.}
+  \label{fig:ex}
+\end{figure}
 
-\todo[inline]{We were unsure what to include in this section because most of our
-  project was correct by construction. Benno and Evan, we will touch base with
-  you about your thoughts as to what to include here. We brainstormed a little
-  bit but couldn't settle on anything that made too much sense. Hoping that the
-  sections above will be suitable for this week's version of the draft. One idea
-  we had would be to show off an example of why the embedded program search is
-  so powerful/useful. This seems to be a somewhat empirical claim that we could
-  spend some time fleshing out.}
+% \newpage
+
+\section{Empirical Evaluation - Ease of Use}
+
+This formalism thus far is lacking ergonomics. The tool is rather impractical
+because the module encoding needs to be written by hand in a coded language that
+relies heavily on numerical indexing. We alleviate this problem with a set of
+programs which both generate the coded modules and the coded queries in
+\textsf{TypeCode}. These two programs hook into the Agda Compiler to parse the
+target library files and generate an auxiliary file containing the coded modules
+and queries. The queries will need to be provided as codes in \textsf{TypeCode}.
+The user provides the syntax valid for Agda and the tool generates the element
+of \textsf{TypeCode} corresponding to the type provided by converting its base
+types to the proper references and function arrows to the coded-arrows.
+
+Along with these programs we have created an integration with emacs to run the
+indexing program whenever emacs saves an Agda file. The query executable can be
+called using a shortcut or from the list of available emacs commands. We have
+found that the inclusion of these features made the search procedure practical
+for the end-user.
 
 \section{Related Work}
-
-\todo[inline]{Edit: Make changes as Benno suggested}
 
@@
 functional properties or examples to help guide the search. Ours is the first
-tool to provide this functionality.
+tool to provide this functionality in a general framework. It is also the first,
+to our knowledge, to formalize the task of type-directed search using decidable
+properties of types.
 
-\textit{draft note: make sure to look into the search programs for other languages.}
+\paragraph{Typing Universes} There has been much work on universes for generic
+programming. Benke et al. create a collection of category-inspired universes
+specifically with the aim of formalizing past work on generic programming in the
+vein of Generic Haskell \cite{benke2003universes}. This work differs from ours
+in its aim, it mainly attempts to automate the writing of functions such as
+generic map or fold which have a common behavior no matter the structure of the
+type. We attempt to be generic over the types themselves but with a more
+concrete aim. This makes this approach to abstract for our use. Another line of
+work aims at formalizing type theory in type theory using quotient inductive
+types\cite{altenkirch2016type}. This work is closer to our own in that they
+encode their universe mutually in a similar way to our encoding. However this
+work aims at verifying type theory within itself and makes heavy use of Homotopy
+Type Theory\cite{HoTT}. The formalism itself is not quite adequate for our
+purposes as it lacks an mechanism for interpreting its type encodings.
 
-\paragraph{Program Synthesis} As was mentioned in Section \# above, program
-search can be viewed as a special case of the more general program synthesis
-problem. It seems hopeful that program search can be later used as a tool in
-program synthesis by treating it as an atomic operation for larger synthesis
-procedures.
+\section{Future Work}
 
-\paragraph{Type Isomorphism} Isomorphisms over types acts as the equivalence
-relation that we use for search in our approach. In reality our approach defines
-a slightly more robust isomorphism relation over types \textit{and} examples.
-This is slightly more powerful than just the types themselves. In future work we
-will look at the theoretical basis for this.
+\subsection{Ergonomic Improvements}
+
+Our toolchain for generating a module from Agda files will always be apt for
+improvement. Ideally, there will be a full integration with the Agda compiler's
+interactive mode. This will speed up the generation program as well as eliminate
+the need for separate executables.
+
+Supporting the generation of first class modules from multiple files is also a
+near term improvement that is technically not too difficult. The name spaces
+will need to be qualified and import conflicts resolved. This will open up the
+practical use of our tool for searching over a large repository of code such as
+the Agda Standard Library.\cite{bove2009brief}
+
+\subsection{Universe for Fully Dependent Types}
+
+One of the items lacking in this project was a successful encoding of a universe
+for fully dependent types. There was a prototype produced that had some
+deficiencies preventing it from having a sound interpretation function. This
+will be the first follow-up work to be conducted. Completing this universe will
+let the search procedure handle example-based searches such as was show in the
+overview section. This need not stop at examples however. Any decidable property
+over terms/functions can be used in the search process. Thus this formalism will
+be much stronger than any program search tools currently available for
+functional languages.
+
+\subsection{Program Search as Synthesis}
+
+A key realization motivating this line of research is the relation of program
+search to the more general problem of program synthesis. Program Synthesis is
+the task of generating a program which satisfies some specification. Types
+provide one such specification. If the synthesis task is then reduced to
+producing some term of a given query type, the synthesis procedure must produce
+an abstract syntax tree that typechecks to the query type. If we restrict the
+solution to have an AST of size 1 it essentially forces the synthesis problem to
+become program search.
+
+This is not a particularly profound idea. However, when considering program
+search as foundational operation in synthesis procedures it becomes more
+motivating. This was the reasoning behind formalizing search within Agda as
+opposed to as a separate tool. With a function we can compose search with other
+functions to build a complex framework for synthesis over dependent types. This
+is our main line of future work in this area.
 
@@
 
+We have developed a formalism for and implemented the foundations of a search
+tool over Agda codebases. This tool allows Agda programmers to search modules
+for terms that match a user-specified type. It can be readily expanded to handle
+Agda's full type system. Following this expansion, a series of work will be
+opened up that will both enable Agda programmers and eventually lead to
+development of dependently-typed synthesis.
+
+\newpage
 %%
@@
 %% consistent spelling of the heading.
-\begin{acks}
-TBD
-\end{acks}
 

### 2020-05-06 paper/paper.tex
big update

@@
 undecidability\cite{HoTT}. The field is currently studying certain decidable
-subsets of the problem thus we opt to maximize flexibility by parameterizing our
-methods over a provided decidable type equality. The problem of type-directed
-program search then just becomes the act of running the decision procedure for
-this equality over the coded-types in our module. Thus, a searchable universe
-can be defined:
+subsets of the problem, thus we opt to maximize flexibility by parameterizing
+our methods over a provided decidable type equality. The problem of
+type-directed program search then just becomes the act of running the decision
+procedure for this equality over the coded types in our module. Thus, a
+searchable universe can be defined:
 

### 2020-05-06 paper/paper.tex
final commit

@@
 the entry may depend on it. The entries in the module will be similar to that in
-the example module in Section 4. In the same way the dependent pair
+the example module in Section 4. This is in the same way the dependent pair
 did in the previous Module definition.
@@
 
-To make all of this a bit more concrete, Figure \ref{fig:ex} is an example of a
-module encoded into this typing universe. Here each entry is preceded by its
-type in the coded type system. Note that the references properly index the types
-of their functions or terms.
+More concretely, Figure \ref{fig:ex} is an example of a module encoded into this
+typing universe. Here each entry is preceded by its type in the coded type
+system. Note that the references properly index the types of their functions or
+terms.
 
@@
 
-This formalism thus far is lacking ergonomics. The tool is rather impractical
+This formalism thus far is lacking in ergonomics. The tool is rather impractical
 because the module encoding needs to be written by hand in a coded language that
@@
 and queries. The queries will need to be provided as codes in \textsf{TypeCode}.
-The user provides the syntax valid for Agda and the tool generates the element
-of \textsf{TypeCode} corresponding to the type provided by converting its base
-types to the proper references and function arrows to the coded-arrows.
+The user provides the syntax valid for Agda and the query tool generates the
+element of \textsf{TypeCode} corresponding to the type provided by converting
+its base types to the proper references and function arrows to the coded-arrows.
 
-Along with these programs we have created an integration with emacs to run the
-indexing program whenever emacs saves an Agda file. The query executable can be
+Along with these programs we have created an integration with Emacs to run the
+indexing program whenever Emacs saves an Agda file. The query executable can be
 called using a shortcut or from the list of available emacs commands. We have
-found that the inclusion of these features made the search procedure practical
+found that the inclusion of these features makes the search procedure practical
 for the end-user.
@@
 type. We attempt to be generic over the types themselves but with a more
-concrete aim. This makes this approach to abstract for our use. Another line of
-work aims at formalizing type theory in type theory using quotient inductive
-types\cite{altenkirch2016type}. This work is closer to our own in that they
-encode their universe mutually in a similar way to our encoding. However this
-work aims at verifying type theory within itself and makes heavy use of Homotopy
-Type Theory\cite{HoTT}. The formalism itself is not quite adequate for our
-purposes as it lacks an mechanism for interpreting its type encodings.
+concrete aim. This approach is too abstract for our use. Another line of work
+aims at formalizing type theory in type theory using quotient inductive
+types\cite{altenkirch2016type}. This work is closer to our own in that it
+encodes the universe mutually. However this work aims at verifying type theory
+within itself and makes heavy use of Homotopy Type Theory\cite{HoTT}. The
+formalism itself is not quite adequate for our purposes as it lacks a mechanism
+for interpreting its type encodings.
 
@@
 Supporting the generation of first class modules from multiple files is also a
-near term improvement that is technically not too difficult. The name spaces
+near-term improvement that is technically not too difficult. The namespaces
 will need to be qualified and import conflicts resolved. This will open up the
 practical use of our tool for searching over a large repository of code such as
-the Agda Standard Library.\cite{bove2009brief}
+the Agda Standard Library\cite{bove2009brief}.
 
@@
 will be the first follow-up work to be conducted. Completing this universe will
-let the search procedure handle example-based searches such as was show in the
-overview section. This need not stop at examples however. Any decidable property
-over terms/functions can be used in the search process. Thus this formalism will
+let the search procedure handle example-based searches such as was shown in the
+overview section. This need not stop at examples. Any decidable property over
+terms or functions can be used in the search process. Thus this formalism will
 be much stronger than any program search tools currently available for
@@
 A key realization motivating this line of research is the relation of program
-search to the more general problem of program synthesis. Program Synthesis is
+search to the more general problem of Program Synthesis. Program Synthesis is
 the task of generating a program which satisfies some specification. Types
@@
 This is not a particularly profound idea. However, when considering program
-search as foundational operation in synthesis procedures it becomes more
-motivating. This was the reasoning behind formalizing search within Agda as
-opposed to as a separate tool. With a function we can compose search with other
+search as a foundational operation in synthesis procedures it becomes more
+interesting. This was the reasoning behind formalizing search within Agda as
+opposed to in a separate tool. With a function we can compose search with other
 functions to build a complex framework for synthesis over dependent types. This
-is our main line of future work in this area.
+is our main line of future work.
 
